---
title: "FA-4 R 6.1"
author: "Espiritu, Joseph Raphael M."
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: true
  rmarkdown: github_document
  word_document: default
  pdf_document: default
---
### **Exercises 6.1**

#### ***5.*** A geospatial analysis system has four sensors supplying images. The percent- age of images supplied by each sensor and the percentage of images relevant to a query are shown in the following table.

``` {r echo= FALSE}
library(gt) 
# Create the data frame
sensor_data <- data.frame(
  Sensor = c(1, 2, 3, 4),
  Images.Supplied = c(15, 20, 25, 40),
  Relevant.Images = c(50, 60, 80, 85)
)

# Generate a gt table
sensor_data %>%
  gt() %>%
  tab_header(
    title = "Sensor Data",
    subtitle = "Percentage of Images Supplied and Relevant Images"
  ) %>%
  fmt_percent(
    columns = c(Images.Supplied, Relevant.Images),  # Format percentages
    scale_values = FALSE,
    decimals = 0
  ) %>%
  cols_label(
    Sensor = "Sensor",
    Images.Supplied = "Images Supplied",
    Relevant.Images = "Relevant Images"
  )
```

What is the overall Percentage of the Relevant Images?

This is one Equation to solve it

\[
P_{\text{overall}} = \frac{\sum (P_{\text{supplied}, i} \times P_{\text{relevant}, i})}{\sum P_{\text{supplied}, i}}
\]

```{r}
percentSupplied <- c(15, 20, 25, 40)
percentRelevant <- c(50, 60, 80, 85)

overallPercentage <- sum(percentSupplied * percentRelevant) / sum(percentSupplied)

cat("Overall percentage of the Relevant images:", overallPercentage, "%\n")
```
#### ***6.*** A fair coin is tossed twice.
Let E be the event that both tosses have the same outcome, that is, E1 = (HH, TT).
Let E2 be the event that the first toss is a head, that is, E2 = (HH, HT).
Let E3 be the event that the second toss is a head, that is, E3 = (TH, HH).
Show that E1, E2, and E3 are pairwise independent but not mutually independent.

``` {r echo = FALSE}
# Create the matrix
coinToss <- matrix(c("HH", "TH", "HT", "TT"), nrow = 2, ncol = 2)
rownames(coinToss) <- c("Heads", "Tails")
colnames(coinToss) <- c("Heads", "Tails")

# Convert to data frame for gt
coinToss_df <- as.data.frame(coinToss)
colnames(coinToss_df) <- c("Heads", "Tails")
coinToss_df <- cbind(Result = c("Heads", "Tails"), coinToss_df)

# Generate a gt table
coinToss_df %>%
  gt() %>%
  tab_header(
    title = "Coin Toss Outcomes",
  )
``` 

\[
(E_1) = \{ \text{HH}, \text{TT} \} \\
\text{E1 is the Event both results are the Same} \\
(E_2) = \{ \text{HH}, \text{HT} \} \\
\text{E2 is the Event first results are the Heads} \\
(E_3) = \{ \text{HH}, \text{TH} \} \\
\text{E3 is the Event second results are the Heads} \\
P(E_1) = P(E_2) = P(E_3) = \frac{1}{2} \\
\]

#### To show PAIRWISE INDEPENDENCE we can just show their Union/Probabilities Together

\[
P(E_i \cap E_j) = P(E_i) \times P(E_j)
\]

For \(E_1\) and \(E_2\):

\[
P(E_1 \cap E_2) = P(\text{HH}) = \frac{1}{4}
\]
\[
P(E_1) \times P(E_2) = \frac{1}{2} \times \frac{1}{2} = \frac{1}{4}
\]

Similarly:

\[
P(E_1 \cap E_3) = P(\text{HH}) = \frac{1}{4}
\]
\[
P(E_1) \times P(E_3) = \frac{1}{2} \times \frac{1}{2} = \frac{1}{4}
\]
\[
P(E_2 \cap E_3) = P(\text{HH}) = \frac{1}{4}
\]
\[
P(E_2) \times P(E_3) = \frac{1}{2} \times \frac{1}{2} = \frac{1}{4}
\]

#### These prove their PAIRWISE INDEPENDENCE, but now to show they are NOT MUTUALLY INDEPENDENT
#### We can apply the same principle of multiplying the probabilities and their unions, and we will see the problem


\[
P(E_1 \cap E_2 \cap E_3) = P(E_1) \times P(E_2) \times P(E_3)
\]

\[
P(E_1 \cap E_2 \cap E_3) = P(\text{HH}) = \frac{1}{4}
\]

\[
P(E_1) \times P(E_2) \times P(E_3) = \frac{1}{2} \times \frac{1}{2} \times \frac{1}{2} = \frac{1}{8}
\]

#### Hence, they are not mutually independent because \( E_2 \) and \( E_3 \) themselves don't carry enough information to help form the probability of \( E_1 \).  

#### However, having the union of \( E_2 \) and \( E_3 \) changes the probability of \( E_1 \) happening, making it guaranteed.





















