---
title: "FA-E1 R 7.1"
author: "Espiritu, Joseph Raphael M."
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: true
  word_document: default
  rmarkdown: github_document
  pdf_document: default
---

---

## Problem Statement 1:

A **malicious spyware** can infect a computer system through either the **Internet** or through **email**. The probabilities for the **entry points** and the **detection rates** are as follows:

- **70%** of spyware infections occur through the **Internet**.
- **30%** of spyware infections occur through **Email**.

If the spyware enters via the **Internet**, the anti-virus detector will detect it with a probability of **60%**. If it enters via **Email**, the anti-virus will detect it with a probability of **80%**.

These probabilities can be expressed as:

- **70%** of all spyware comes through the **Internet**: \( P(\text{Internet}) = 0.70 \)
- **30%** of all spyware comes through **Email**: \( P(\text{Email}) = 0.30 \)
- The probability of detection if the spyware is from the **Internet**: \( P(\text{Detected} | \text{Internet}) = 0.60 \)
- The probability of detection if the spyware is from **Email**: \( P(\text{Detected} | \text{Email}) = 0.80 \)
- The probability of undetected spyware from the **Internet**: \( P(\text{Undetected} | \text{Internet}) = 0.40 \)
- The probability of undetected spyware from **Email**: \( P(\text{Undetected} | \text{Email}) = 0.20 \)

### Questions:

(a) What is the probability that this **spyware infects the system**?

(b) If the spyware is **detected**, what is the **probability that it came through the
Internet**?

### Solution

- The probability of detected spyware is: 

\[
P(\text{Detected}) = P(\text{Detected} | \text{Internet}) P(\text{Internet}) + P(\text{Detected} | \text{Email}) P(\text{Email})
\]

``` {r}
# Given Probabilities
P_Internet <- 0.70
P_Email <- 0.30
P_Detected_Internet <- 0.60
P_Detected_Email <- 0.80
P_Undetected_Internet <- 0.40
P_Undetected_Email <- 0.20

# Solving for Detected Spyware
P_Detected <- P_Detected_Internet * P_Internet + P_Detected_Email * P_Email

# Show Value
P_Detected

```

Using **law of total probability**, the probability of receiving **Undetected spyware** is:

\[
P(\text{Undetected}) = P(\text{Undetected} | \text{Internet}) P(\text{Internet}) + P(\text{Undetected} | \text{Email}) P(\text{Email})
\]

Using **Bayes' theorem**, the probability that a **spyware detected from the internet** is:

\[
P(\text{Internet} | \text{Detected}) = \frac{P(\text{Detected} | \text{Internet}) P(\text{Internet})}{P(\text{Detected})}
\]

---

```{r}
# Solving for Spyware Undetected
P_Undetected <- P_Undetected_Internet * P_Internet + P_Undetected_Email * P_Email

# Solving for Spyware detected from Internet
P_Internet_Detected <- P_Detected_Internet * P_Internet / P_Detected

# Show Values
P_Undetected
P_Internet_Detected

```
---

### Thus, **Undetected Spyware appears** `r P_Undetected*100`% and **Detected spyware from the Internet appears** around `r round(P_Internet_Detected * 100, 2)`% in any computer system. 

## Problem Statement 2:

Of the **emails you receive, 20% are spam** on average. Your **spam filter** is able to
detect **90%** of them but also **misclassifies as spam 15% of the genuine emails**.

These probabilities can be expressed as:

- Probability that an email is spam: \( P(\text{Spam}) = 0.20 \)
- Probability that an email is legitimate: \( P(\text{Legitimate}) = 0.80 \)
- Probability that the spam filter detects spam: \( P(\text{Marked} | \text{Spam}) = 0.90 \)
- Probability that the spam filter misclassifies a legitimate email as spam: \( P(\text{Marked} | \text{Legitimate}) = 0.15 \)
- Probability that the spam filter does not mark a spam email as spam: \( P(\text{Not Marked} | \text{Spam}) =  0.10 \)
- Probability that the spam filter does not mark a legitimate email as spam: \( P(\text{Not Marked} | \text{Legitimate}) = 0.85 \)

### Questions:

(a) If an **email arrives** and is **marked spam**, what is the **probability that it really
is spam**?
(b) If an **email arrives** and is **not marked spam**, what is the **probability that it
is legitimate**?

### Solution

- The probability of Marked Spam is: 

\[
P(\text{Marked}) = P(\text{Marked} | \text{Spam}) P(\text{Spam}) + P(\text{Marked} | \text{Legitimate}) P(\text{Legitimate})
\]

- The probability of Not Marked Spam is: 

\[
P(\text{Not Marked}) = P(\text{Not Marked} | \text{Spam}) P(\text{Spam}) + P(\text{Not Marked} | \text{Legitimate}) P(\text{Legitimate})
\]

---

``` {r}
# Given Probabilities
P_Spam <- 0.20
P_Legitimate <- 0.80
P_Marked_Spam <- 0.90
P_Marked_Legitimate <- 0.15
P_NotMarked_Spam <- 0.10
P_NotMarked_Legitimate <- 0.85

# Solving for Marked
P_Marked <- P_Marked_Spam * P_Spam + P_Marked_Legitimate * P_Legitimate

# Solving for Not Marked
P_NotMarked <- P_NotMarked_Spam * P_Spam + P_NotMarked_Legitimate * P_Legitimate

# Show Value
P_Marked
P_NotMarked

```

---

Therefore,

- Probability for anything to be Marked is: \( P(\text{Marked}) = 0.3 \)
- Probability for anything to be Not Marked: \( P(\text{NotMarked}) = 0.70 \)

Using **Bayes' theorem** for both questions we have the following equations,

\[
P(\text{Spam} | \text{Marked}) = \frac{P(\text{Marked} | \text{Spam}) \cdot P(\text{Spam})}{P(\text{Marked})}
\]
\[
P(\text{Legitimate} | \text{Not Marked}) = \frac{P(\text{Not Marked} | \text{Legitimate}) \cdot P(\text{Legitimate})}{P(\text{Not Marked})}
\]

---

```{r}
# Solving for Spam when its Marked
P_Spam_Marked <- P_Marked_Spam * P_Spam / P_Marked

# Solving for Legitimate when its Not Marked
P_Legitimate_NotMarked <- P_NotMarked_Legitimate * P_Legitimate / P_NotMarked

# Show Values
P_Spam_Marked
P_Legitimate_NotMarked

```
---

### Thus, **if an email arrives**, the **probability it is spam** is about `r P_Spam_Marked * 100`%, and the **probability that a legitimate email** is **not marked as spam** is around `r round(P_Legitimate_NotMarked * 100, 2)`%.

---
